Action Groupoids and Fractional Types

### 2016-02-16 frac.tex
organizing the flow in the paper better and starting to formalize things a bit

\documentclass{article}
\usepackage{fullpage}
\usepackage{bbold}
\usepackage{url}

\newcommand{\inl}[1]{\textsf{inl}(#1)}
\newcommand{\inr}[1]{\textsf{inr}(#1)}
\newcommand{\zt}{\mathbb{0}}
\newcommand{\ot}{\mathbb{1}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\fv}[2]{\langle #1 ~|~ #2 \rangle}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Zn}{\mathbb{Z}_n}
\newcommand{\Zvn}{\mathbb{Z}^v_n}
\newcommand{\pt}[2]{#1_{\bullet}^{#2}}
\newcommand{\cycle}{\textsf{cycle}}

@@

Our starting point is $\Pi$:
\begin{itemize}
\item We have finite types $\zt$, $\ot$, $\tau_1\oplus\tau_2$, and $\tau_1\otimes\tau_2$
\item The types have points $()$, $\inl{v}$, $\inr{v}$, and $(v_1,v_2)$
\item A type $\tau$ has $|\tau|$ points
\item We have combinators $c : \tau_1\leftrightarrow\tau_2$ between the types which witness type isomorphisms and which correspond to the axioms of commutative rigs
\item If we have combinators $c_1, c_2 : \tau_1\leftrightarrow\tau_2$, we have level-2 combinators
$\alpha : c_1 \Leftrightarrow c_2$ which look quite messy but they are essentially witnesses of
the coherence conditions for rig groupoids.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Generalization}

Now we do the following:
\begin{itemize}
\item We generalize the syntax of types to include fractional types $\tau_1/\tau_2$ 

>>>>

\documentclass{article}
\usepackage{fullpage}
\usepackage{bbold}
\usepackage{url}
\usepackage{xcolor}
\usepackage{amssymb}

\newcommand{\alt}{~|~}
\newcommand{\inl}[1]{\textsf{inl}(#1)}
\newcommand{\inr}[1]{\textsf{inr}(#1)}
\newcommand{\zt}{\mathbb{0}}
\newcommand{\ot}{\mathbb{1}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\fv}[2]{\langle #1 ~|~ #2 \rangle}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Zn}{\mathbb{Z}_n}
\newcommand{\Zvn}{\mathbb{Z}^v_n}
\newcommand{\pt}[2]{#1_{\bullet}^{#2}}
\newcommand{\cycle}{\textsf{cycle}}
\newcommand{\twod}{\mathbb{T}}
\newcommand{\fract}[2]{\fcolorbox{black}{white}{$#1$}\fcolorbox{black}{gray!20}{$#2$}}

@@

Our starting point is $\Pi$:
\begin{itemize}
\item We have finite types $\zt$, $\ot$, $\tau_1\oplus\tau_2$, and $\tau_1\otimes\tau_2$
\item The types have points $()$, $\inl{v}$, $\inr{v}$, and $(v_1,v_2)$
\item A type $\tau$ has $|\tau|$ points:
\[\begin{array}{rcl}
|\zt| &=& 0 \\
|\ot| &=& 1 \\
|\tau_1 \oplus \tau_2| &=& |\tau_1| + |\tau_2| \\
|\tau_1 \otimes \tau_2| &=& |\tau_1| * |\tau_2| 
\end{array}\]
\item We have combinators $c : \tau_1\leftrightarrow\tau_2$ between
  the types which witness type isomorphisms and which correspond to
  the axioms of commutative rigs
\item If we have combinators $c_1, c_2 : \tau_1\leftrightarrow\tau_2$,
  we have level-2 combinators $\alpha : c_1 \Leftrightarrow c_2$ which
  are (quite messy) equivalences of isomorphisms, and which happen to
  correspond to the coherence conditions for rig groupoids.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{2D-types: Intuition}

In HoTT, types are groupoids. One can visualize a type as a collection
of points connected by paths. The paths can be arbitrary subject to
three conditions: there is an associative operation which composes
paths; there is an identity path; and every path has an inverse.  This
most general form is difficult to capture structurally and
computationally but there are some interesting special cases, one of
which we explore in detail in this paper. The special case we study in
detail is that of an \emph{action groupoid} $S \rtimes \G$. This is a
groupoid in which the paths are generated by group.

Give lots of examples of action groupoids. Explain cardinality.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{2D-types: Syntax}

\begin{itemize}
\item We generalize the syntax of types to include fractional types
  $\tau_1/\tau_2$ which we will write as $\fract{\tau_1}{\tau_2}$ with
  the numerator in white and the denominator in grey. The idea will be
  that $\tau_1$ denotes a set $S$ and $\tau_2$ denotes a group $\G$
  and that $\fract{\tau_1}{\tau_2}$ denotes the action groupoid
  $S \rtimes \G$.

\item We actually have two levels of types:
\[\begin{array}{rcl} 
\tau &::=& \zt \alt \ot \alt \tau_1\oplus\tau_2 \alt \tau_1\otimes\tau_2 \\
\twod &::=& \fract{\tau_1}{\tau_2} \alt \twod_1 \boxplus \twod_2
            \alt \twod_1 \boxtimes \twod_2  
\end{array}\]
The $\tau$ level describes plain sets. The $\twod$ level describes
``two-dimensional types'' which will be explained in more detail
later. 
\item Note that 2D types are closed under sums and products but
  \emph{not} under ``division.'' In other words, we cannot form types
  $\fract{\fract{\tau_1}{\tau_2}}{\fract{\tau_3}{\tau_4}}$. This is
  why we call our types 2D as we are restricted to two levels.

\item Pointed types\ldots
